\documentclass[11pt]{letter}
\usepackage[utf8]{inputenc}
\usepackage{geometry}
\geometry{a4paper, margin=1in}
\usepackage{hyperref}

\signature{David Ahmann\\Independent Researcher\\Toronto, Canada\\\texttt{dahmann@lumyn.cc}}

\begin{document}

\begin{letter}{Editors-in-Chief\\Journal of Machine Learning Research (JMLR)}

\opening{Dear Editors,}

I am pleased to submit the manuscript entitled ``ALFM-BEM: Bidirectional Experience Memory for Continuous Learning in Foundation Model Deployments'' for consideration as a publication in the Journal of Machine Learning Research.

\section*{Summary}

Foundation models are typically deployed as frozen artifacts, creating a fundamental gap: they cannot learn from their deployment experiences. This paper introduces ALFM-BEM, a unified wrapper architecture that enables continuous learning without modifying backbone weights. Key contributions include:
\begin{itemize}
    \item \textbf{Bidirectional Experience Memory (BEM)}: A unified memory architecture where experiences exist on a continuous outcome spectrum, providing risk signals, success patterns, and out-of-distribution detection as emergent properties of a single structure.
    \item \textbf{Consensus Engine with Query Action}: An extension to selective prediction that transforms passive abstention into active learning.
    \item \textbf{Bounded Adapters}: Continuous improvement with provable stability guarantees (Proposition 4.4).
    \item \textbf{Empirical Validation}: Experiments demonstrate effective failure retrieval (F1 $\approx$ 0.59), strong OOD detection (AUC $>$ 0.99 for clustered patterns), and unique success pattern retrieval (rate $\approx$ 0.70) that baselines cannot provide.
\end{itemize}

\section*{Previous Publications}

This manuscript has not been published previously at any workshop or conference. This is an original submission.

\section*{Co-Author Consent}

I am the sole author of this manuscript and consent to its review by JMLR.

\section*{Conflicts of Interest}

I declare no conflicts of interest with any JMLR action editors or reviewers. I have no financial relationships with entities that could be perceived to influence this work.

\section*{Funding Disclosure}

This research was conducted independently without external funding. No grants, stipends, donations, or third-party support were received for any aspect of this work, including data collection, computing resources, or development.

\section*{Suggested Action Editors}

Based on expertise in memory-augmented neural networks, continual learning, and foundation model deployment:
\begin{enumerate}
    \item \textbf{Finale Doshi-Velez} -- Expertise in interpretable ML and safe deployment
    \item \textbf{Been Kim} -- Expertise in interpretability and model understanding
    \item \textbf{Percy Liang} -- Expertise in foundation models and robustness
    \item \textbf{Jacob Steinhardt} -- Expertise in ML safety and distribution shift
    \item \textbf{Zachary Lipton} -- Expertise in deployment and calibration
\end{enumerate}

\section*{Suggested Reviewers}

\begin{enumerate}
    \item \textbf{Yonatan Geifman} (Technion) -- Expert on selective prediction
    \item \textbf{David Rolnick} (Mila) -- Expert on experience replay and continual learning
    \item \textbf{Dan Hendrycks} (Center for AI Safety) -- Expert on OOD detection
    \item \textbf{Yaniv Taigman} (Meta) -- Expert on memory networks
    \item \textbf{Charles Packer} (UC Berkeley) -- Expert on memory systems for LLMs (MemGPT)
\end{enumerate}

\section*{Keywords}

Continual learning, experience memory, out-of-distribution detection, foundation models, selective prediction, bounded adapters, retrieval-augmented systems

\section*{Paper Length Justification}

The manuscript is approximately 15 pages (main text) plus appendices, well within JMLR guidelines.

\closing{Sincerely,}

\end{letter}
\end{document}
